---
title: "4_analysis_and_generating_figures"
output: html_document
---

#load library
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bigleaf)
library(tidyverse)
library(RNetCDF)
library(fields)
library(akima)
library(RColorBrewer)
library(maps)

```


# list of models
```{r}
### list of models we use:
list_models_common <- c('ACCESS-CM2', 'AWI-CM-1-1-MR','CanESM5-CanOE','CESM2','CMCC-CM2-SR5','CMCC-ESM2',
												'CNRM-CM6-1',	'CNRM-ESM2-1','E3SM-1-1','FGOALS-f3-L','FGOALS-g3',	'FIO-ESM-2-0', 'GFDL-ESM4',
												'HadGEM3-GC31-LL','HadGEM3-GC31-MM','INM-CM4-8','INM-CM5-0',	'IPSL-CM6A-LR','KACE-1-0-G',
												'MIROC6','MIROC-ES2L','MPI-ESM1-2-LR','MRI-ESM2-0','NorESM2-MM','TaiESM1','UKESM1-0-LL')
```



#load data
```{r}
load("./analysis_ready_data/AI_climate_historical_2x2_allmodels.RData")
load("./analysis_ready_data/NAI_climate_historical_2x2_allmodels.RData")
load("./analysis_ready_data/pr_climate_historical_2x2_allmodels.RData")
load("./analysis_ready_data/PE_climate_historical_2x2_allmodels.RData")
load("./analysis_ready_data/SFE_climate_historical_2x2_allmodels.RData")

load("./analysis_ready_data/AI_climate_ssp585_2x2_allmodels.RData")
load("./analysis_ready_data/NAI_climate_ssp585_2x2_allmodels.RData")
load("./analysis_ready_data/pr_climate_ssp585_2x2_allmodels.RData")
load("./analysis_ready_data/SFE_climate_ssp585_2x2_allmodels.RData")
load("./analysis_ready_data/PE_climate_ssp585_2x2_allmodels.RData")

load("./analysis_ready_data/mask_2x2_NAs.RData")
load("./analysis_ready_data/mask_2x2.RData")
```

#cos weigth
```{r}
## 2X2 cos weight: this can be used to calculate global mean
lat <- seq(from = -90, to = 90, by = 2)
weight <- cos(lat  * (pi/180))
w <- t(array(weight,dim = c(91,181)))
weight <- w/sum(w)
image.plot(weight)

```

### From now on, you need to analyze data on your own
example 1: AI and NAI maps with dryland classification line (0.65 for AI, 1 for NAI)
example 2: comparison between AI and NAI 
example 3: Future dryland fraction - historical dryland fraction (for AI and NAI. then compare. statistical analysis)

I recommend referring to Berg & McColl (2021) and their r code.
paper: https://doi.org/10.1038/s41558-021-01007-8
code: https://doi.org/10.5281/zenodo.4490414


AI dryland
```{r}
## historical
mean_AI <- apply(AI_climate_historical_2x2_allmodels, c(1, 2), mean, na.rm=T)

## Lon/Lat and mask on 2x2:
lat <- mask_2x2$y
lon <-  mask_2x2$x
lowlat <- min(which(lat > -60));highlat <- min(which(lat > 80)); mask <- mask_2x2_NAs

##
lowlat <- min(which(lat > -60)); 
highlat <- min(which(lat > 80))

bob <- mean_AI; bob[bob >4 ] <- 4;
image.plot(lon-180, lat[lowlat:highlat], bob[ c( (length(lon)/2+1):length(lon),
1:(length(lon)/2)),lowlat:highlat],  zlim=c(0,4),col=brewer.pal(8, "YlGnBu"), breaks=seq(0,4,by=0.5),
xlab="", ylab="", xaxt="n", yaxt="n", axis.args=list(cex.axis=1.5), cex.main=1.9,main="Mean AI and drylands");
map(add=T,interior=F)
bob[is.na(bob)==T] <- 1000
contour(seq(-180,180, length.out=180), seq(lat[lowlat],lat[highlat],length.out=length(lowlat:highlat)),
bob[ c( (length(lon)/2+1):length(lon), 1:(length(lon)/2)),lowlat:highlat],levels=c(0.5,0.65),
 col=c("orange","darkred"), lwd=1, add=T, drawlabels=F); 
map(add=T,interior=F)
#legend("bottomleft", col=c("orange", "darkred"), c("AI=0.5", "AI=0.65"), bty="n", lwd=1.2, cex=1.6)
mtext("a", side=3, adj=0, line=1, cex=1.5, font=2)


## future
mean_AI <- apply(AI_climate_ssp585_2x2_allmodels, c(1, 2), mean, na.rm=T)

## Lon/Lat and mask on 2x2:
lat <- mask_2x2$y
lon <-  mask_2x2$x
lowlat <- min(which(lat > -60));highlat <- min(which(lat > 80)); mask <- mask_2x2_NAs

##
lowlat <- min(which(lat > -60)); 
highlat <- min(which(lat > 80))

bob <- mean_AI; bob[bob >4 ] <- 4;
image.plot(lon-180, lat[lowlat:highlat], bob[ c( (length(lon)/2+1):length(lon),
1:(length(lon)/2)),lowlat:highlat],  zlim=c(0,4),col=brewer.pal(8, "YlGnBu"), breaks=seq(0,4,by=0.5),
xlab="", ylab="", xaxt="n", yaxt="n", axis.args=list(cex.axis=1.5), cex.main=1.9,main="Mean AI and drylands");
map(add=T,interior=F)
bob[is.na(bob)==T] <- 1000
contour(seq(-180,180, length.out=180), seq(lat[lowlat],lat[highlat],length.out=length(lowlat:highlat)),
bob[ c( (length(lon)/2+1):length(lon), 1:(length(lon)/2)),lowlat:highlat],levels=c(0.5,0.65),
 col=c("orange","darkred"), lwd=1, add=T, drawlabels=F); 
map(add=T,interior=F)
#legend("bottomleft", col=c("orange", "darkred"), c("AI=0.5", "AI=0.65"), bty="n", lwd=1.2, cex=1.6)
mtext("a", side=3, adj=0, line=1, cex=1.5, font=2)

```


NAI dryland
```{r}
mean_NAI <- apply(NAI_climate_historical_2x2_allmodels, c(1, 2), mean, na.rm=T)

## Lon/Lat and mask on 2x2:
lat <- mask_2x2$y
lon <-  mask_2x2$x
lowlat <- min(which(lat > -60));highlat <- min(which(lat > 80)); mask <- mask_2x2_NAs

##
lowlat <- min(which(lat > -60)); 
highlat <- min(which(lat > 80))

bob <- mean_NAI; bob[bob >4 ] <- 4;
image.plot(lon-180, lat[lowlat:highlat], bob[ c( (length(lon)/2+1):length(lon),
1:(length(lon)/2)),lowlat:highlat],  zlim=c(0,4),col=brewer.pal(8, "YlGnBu"), breaks=seq(0,4,by=0.5),
xlab="", ylab="", xaxt="n", yaxt="n", axis.args=list(cex.axis=1.5), cex.main=1.9,main="Mean NAI and drylands");
map(add=T,interior=F)
bob[is.na(bob)==T] <- 1000
contour(seq(-180,180, length.out=180), seq(lat[lowlat],lat[highlat],length.out=length(lowlat:highlat)),
bob[ c( (length(lon)/2+1):length(lon), 1:(length(lon)/2)),lowlat:highlat],levels=c(1),
 col=c("darkred"), lwd=1, add=T, drawlabels=F); 
map(add=T,interior=F)
#legend("bottomleft", col=c("darkred"), c("NAI=1"), bty="n", lwd=1.1, cex=1.6)
mtext("a", side=3, adj=0, line=1, cex=1.5, font=2)

### future
mean_NAI <- apply(NAI_climate_ssp585_2x2_allmodels, c(1, 2), mean, na.rm=T)

## Lon/Lat and mask on 2x2:
lat <- mask_2x2$y
lon <-  mask_2x2$x
lowlat <- min(which(lat > -60));highlat <- min(which(lat > 80)); mask <- mask_2x2_NAs

##
lowlat <- min(which(lat > -60)); 
highlat <- min(which(lat > 80))

bob <- mean_NAI; bob[bob >4 ] <- 4;
image.plot(lon-180, lat[lowlat:highlat], bob[ c( (length(lon)/2+1):length(lon),
1:(length(lon)/2)),lowlat:highlat],  zlim=c(0,4),col=brewer.pal(8, "YlGnBu"), breaks=seq(0,4,by=0.5),
xlab="", ylab="", xaxt="n", yaxt="n", axis.args=list(cex.axis=1.5), cex.main=1.9,main="Mean NAI and drylands");
map(add=T,interior=F)
bob[is.na(bob)==T] <- 1000
contour(seq(-180,180, length.out=180), seq(lat[lowlat],lat[highlat],length.out=length(lowlat:highlat)),
bob[ c( (length(lon)/2+1):length(lon), 1:(length(lon)/2)),lowlat:highlat],levels=c(1),
 col=c("darkred"), lwd=1, add=T, drawlabels=F); 
map(add=T,interior=F)
#legend("bottomleft", col=c("darkred"), c("NAI=1"), bty="n", lwd=1.1, cex=1.6)
mtext("a", side=3, adj=0, line=1, cex=1.5, font=2)


```


```{r}

dryland_fraction_AI <- data.frame(list_models = list_models_common,
																	historical = rep(NA, length(list_models_common)),
																	future = rep(NA, length(list_models_common))
																	)

dryland_fraction_NAI <- data.frame(list_models = list_models_common,
																	 historical = rep(NA, length(list_models_common)),
																	 future = rep(NA, length(list_models_common))
																	)
	
for(i in c(1:length(list_models_common))){
	dryland_fraction_AI$historical[i] <-
		sum((mask_2x2_NAs*weight)[which(AI_climate_historical_2x2_allmodels[,,i] < 0.65)], na.rm=T)/
		sum((mask_2x2_NAs*weight), na.rm=T)* 100
	dryland_fraction_AI$future[i] <-
		sum((mask_2x2_NAs*weight)[which(AI_climate_ssp585_2x2_allmodels[,,i] < 0.65)], na.rm=T)/
		sum((mask_2x2_NAs*weight), na.rm=T)* 100
	
	dryland_fraction_NAI$historical[i] <-
		sum((mask_2x2_NAs*weight)[which(NAI_climate_historical_2x2_allmodels[,,i] <= 1)], na.rm=T)/
		sum((mask_2x2_NAs*weight), na.rm=T)* 100
	dryland_fraction_NAI$future[i] <-
		sum((mask_2x2_NAs*weight)[which(NAI_climate_ssp585_2x2_allmodels[,,i] <= 1)], na.rm=T)/
		sum((mask_2x2_NAs*weight), na.rm=T)* 100
	
}

mean(dryland_fraction_AI$future)
mean(dryland_fraction_AI$historical)	
mean(dryland_fraction_NAI$future)
mean(dryland_fraction_NAI$historical)

mean(dryland_fraction_AI$future) - mean(dryland_fraction_AI$historical)
mean(dryland_fraction_NAI$future)- mean(dryland_fraction_NAI$historical)

```



